---
title: "Exercise #3"
subtitle: "Fortgeschrittene Statistische Software für NF - SS 2022/23"
author: "Pia Baader (12162232)"
date: "`r Sys.Date()`"
---

## General Remarks

-   You can submit your solutions in teams of up to 3 students.
-   Include all your team-member's names and student numbers
    (Martrikelnummern) in the `authors` field.
-   Please use the exercise template document to work on and submit your
    results.
-   Use a level 2 heading for each new exercise and answer each subtask
    next to it's bullet point or use a new level 3 heading if you want.
-   Always render the R code for your solutions and make sure to include
    the resulting data in your rendered document.
    -   Make sure to not print more than 10 rows of data (unless
        specifically instructed to).
-   Always submit both the rendered document(s) as well as your source
    Rmarkdown document. Submit the files separately on moodle, **not**
    as a zip archive.

## Exercise 1: Initializing git (4 Points)

For this whole exercise sheet we will be tracking all our changes to it
in git.

a)  Start by initializing a new R project with git support, called
    `exeRcise-sheet-3`. If you forgot how to do this, you can follow
    this
    [guide](https://malikaihle.github.io/Introduction-RStudio-Git-GitHub/rstudio_project.html).
b)  Commit the files generated by Rstudio.
c)  For all of the following tasks in this exercise sheet we ask you to
    always commit your changes after finishing each subtask e.g. create
    a commit after task *1d*, *1e* etc.

> Note: This applies only to answers that have text or code as their
> answer. If you complete tasks in a different order or forget to commit
> one, this is no problem. If you change your answers you can just
> create multiple commits to track the changes.

d)  Name 2 strengths and 2 weaknesses of git. (Don't forget to create a
    commit after this answer, see *1c*)
    
2 strengths: 

- One strength is that it is possible to see which changes were made and the document cannot be easily deleted. 
- Another advantage: working with a team on a project gets easier with git since everyone can access the document and make changes

2 weaknesses: 

- One weakness is that git can be complicated to use for beginners. 
- Working in larger teams can be challenging since git relies mainly on one repository. 




e)  Knit this exercise sheet. Some new files will automatically be
    generated when knitting the sheet e.g. the HTML page. Ignore these
    files, as we only want to track the source files themselves.

## Exercise 2: Putting your Repository on GitHub (3.5 Points)

For this task you will upload your solution to GitHub.

a)  Create a new repository on GitHub in your account named
    `exeRcise-sheet-3`. Make sure you create a **public repository** so
    we are able to see it for grading. Add the link to the repository
    below:

git remote add origin git@github.com:cansubaader/exeRcise-sheet-3.git
git branch -M main
git push -u origin main 


b)  Push your code to this new repository by copying and executing the
    snippet on github listed under
    `…or push an existing repository from the command line`.
c)  Regularly push your latest changes to GitHub again and especially do
    so when you are finished with this sheet.

## Exercise 3: Baby-Names in Munich (4.5 Points)

Download the latest open datasets on given names ("Vornamen") from the
open data repository of the city of Munich for the years `2022` and
`2021`.

Link: <https://opendata.muenchen.de/dataset/vornamen-von-neugeborenen>

a)  Download the data for both years and track it in git. For small
    datasets like these adding them to git is not a problem.

b)  Load the data for both years into R. Check the type of the count
    variable ("Anzahl") and look into the data to determine why it is
    not numeric? Fix the problem in an appropriate manner, it is OK if
    some of the counts are inaccurate because of this. Explain your
    solution and the repercussions.
    
```{r}
library(tidyverse) 

given_names_2021 <- read.csv("vornamen_2021.csv")

given_names_2022 <- read.csv("open_data_portal_2022.csv")

View(given_names_2021)
View(given_names_2022)
```

```{r}
is.numeric(given_names_2021$Anzahl)
is.numeric(given_names_2022$Anzahl)
```
"Anzahl" is not numeric because some names are given "4 oder weniger", not just in numbers but in letters/ words. 

```{r}
given_names_2021 <- given_names_2021 %>% 
    na_if("4 oder weniger") 

given_names_2022 <- given_names_2022 %>% 
    na_if("4 oder weniger")

given_names_2021 <- coalesce(given_names_2021, 0) 

given_names_2022 <- coalesce(given_names_2022, 0)


head(given_names_2021)
tail(given_names_2021)
 
given_names_2021 %>% 
    arrange(Anzahl)

 
 

 

```

Since "4 oder weniger" has letters and cannot be numeric my idea was to remove the NAs with the result that only the numbers stay in the datasets 

c)  Calculate the total number of babies born in Munich in 2022
    and 2021. Which year had the bigger baby-boom?
    
```{r}
#totat_sum_babies_2022 <- sum(given_names_2022$Anzahl)


#total_sum_babies_2021 <- sum(given_names_2021$Anzahl)
```
```{r}
Vornamen <- given_names_2022 %>% 
    select( Anzahl, Vorname, year) 

Vornamen
```


d)  Add a new column `year` to both datasets which holds the correct
    year for each.
    
```{r}
 

given_names_2021 <- given_names_2021 %>% 
    mutate(year = 2021) 
     

given_names_2022 <- given_names_2022 %>% 
    mutate(year = 2022) 

View(given_names_2021)
View(given_names_2022)
 
```


e)  Combine both datasets into one using `bind_rows()`.

```{r}
library(dplyr)
given_names <- bind_rows(given_names_2021, given_names_2022, by = year)
View(given_names)
```


f)  Combine the counts for same names to determine the most popular
    names across both years. Print out the top 10 names in a nicely
    formatted table for both years. Include a table caption.

```{r}
 

top_ten_names <- given_names %>%
  group_by(Anzahl) %>%
  summarize(total_count = sum(Anzahl)) %>%
  arrange(desc(total_count)) %>%
  top_n(10)

```

## Exercise 4: Chat GPT + apply (3 points)

For this task: Specifically use ChatGPT to solve the task and submit
your prompts in addition to the solution

a)  The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

###Create a sample data frame

tax_data \<- data.frame( Name = c("Munich GmbH", "ABC Inc.", "Backpacks
1980", "Bavarian Circus"), Tax_2019 = c(5000, 4000, 6000, 3500),
Tax_2020 = c(4800, 4200, 5800, 3700), Tax_2021 = c(5200, 3800, 5900,
3400) )

###Calculate column-wise means

column_means \<- lapply(tax_data\[, -1\], 2, mean)

column_means

##Answer: 

The correct function is sapply(). 

Chat GPT gave this reason why the given code didn't work: 

"The lapply() function applies a function to each element of a list or vector and returns a list as the result. In the incorrect code, lapply(tax_data[, -1], 2, mean), the second argument 2 is not a valid argument for lapply(). This results in an error because lapply() expects a function as the second argument, not a numeric value."

##corrected code: 

# Create a sample data frame
tax_data <- data.frame(
  Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
  Tax_2019 = c(5000, 4000, 6000, 3500),
  Tax_2020 = c(4800, 4200, 5800, 3700),
  Tax_2021 = c(5200, 3800, 5900, 3400)
)

# Calculate column-wise means
column_means <- sapply(tax_data[, -1], mean)

# Calculate row-wise means
row_means <- rowMeans(tax_data[, -1])

column_means
row_means

 




b)  Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
    
##Answer: 

"The rapply() function in R is used for recursive application of a function to all elements of a list or nested list structure. It allows you to apply a specified function to each individual element and continues the recursion if the element itself is a list. This makes it ideal for traversing complex nested data structures. The function can be customized to selectively apply the function based on the class of the elements. rapply() simplifies the process of applying a function to all elements of a nested list, saving you from writing explicit looping code." 



#Easy example generated by Chat gpt: 
# Create a nested list structure
my_list <- list(a = 1:3, b = list(c = 4:6, d = 7:9))

# Define a function to calculate the sum of elements
sum_elements <- function(x) {
  sum(x)
}

# Apply the sum_elements function recursively to all elements
result <- rapply(my_list, sum_elements)

# Print the result
print(result)




## Final Note

Make sure to push all your commits and changes to GitHub before
submittining the exercise sheet.


prompts: 
I used the task as prompt: The code below does not work because the wrong apply function has
    been used. Find out which apply function would be correct and why it
    did not work. Correct the code. Also calculate the rowwise means.

Solution: 
"The issue in the provided code lies with the usage of the lapply function. The second argument of lapply is supposed to be the function to be applied to each element of the input object. In this case, 2 is being used as the function, which is incorrect.

To calculate the column-wise means, you can use the colMeans function instead. 

# Create a sample data frame
tax_data <- data.frame(Name = c("Munich GmbH", "ABC Inc.", "Backpacks 1980", "Bavarian Circus"),
                       Tax_2019 = c(5000, 4000, 6000, 3500),
                       Tax_2020 = c(4800, 4200, 5800, 3700),
                       Tax_2021 = c(5200, 3800, 5900, 3400))

# Calculate column-wise means
column_means <- colMeans(tax_data[, -1])

# Calculate row-wise means
row_means <- rowMeans(tax_data[, -1])

# Display the column-wise means
column_means

# Display the row-wise means
row_means

The correct function is `ColMeans` and `rowMeans`. 
"The corrected code uses the colMeans function to calculate the column-wise means and the rowMeans function to calculate the row-wise means. The column-wise means are stored in the column_means variable, and the row-wise means are stored in the row_means variable."

b)  Using ChatGPT try to understand what the rapply() function does.
    Create an easy example with mock data where the function is used and
    explain it in your words.
    
prompt: 
what does the rapply() function do? create an easy example with mock data where the function is used. 
Solution: 

"The rapply() function in R is a recursive version of the base apply() function. It is designed to apply a given function to all elements of a list or nested list structure, including nested lists within lists. It traverses the structure recursively, applying the function to each element."

in my own words: 

code provided by chat gpt: 
# Create a nested list structure
nested_list <- list(
  a = 1,
  b = list(
    c = 2,
    d = list(
      e = 3,
      f = 4
    )
  ),
  g = 5
)

# Define a function to increment each element by 1
increment <- function(x) {
  x + 1
}

# Apply the increment function to all elements of the nested list structure
result <- rapply(nested_list, increment, how = "replace")

# Print the modified nested list structure
print(result)
